---
title: "Reproducible Research Course Project 1"
author: Rong Cai
date: Oct 26, 2016
output: md_document
---

##Reproducible Research Course Project 1

Introduction

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded from the course web site:

Dataset: Activity monitoring data [52K]
The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

##read data from my working directory
```{r}
mydata <- read.csv("activity.csv")
head(mydata)
```
Remove missing step value from mydata for data analysis
```{r}
mydata1 <- mydata[complete.cases(mydata$steps),]
head(mydata1)
```
##What is mean total number of steps taken per day?
Calculate the total number of steps taken per day
```{r}
Tstep <- with(mydata1,aggregate(list(steps=steps), by=list(date=date), sum))
head(Tstep)
```
Plot histogram of the total number of steps taken each day
```{r}
library(ggplot2)
g <- qplot(steps, data=Tstep, geom="histogram",bins=20)
g+xlab("Total Steps")+ylab("Frequency")+ggtitle("Total Number of Steps Taken Each Day")
```
Calculate and report the mean and median of the total number of steps taken per day
```{r}
##mean
MNstep <- mean(Tstep$steps)
MNstep
##median
MDstep <- median(Tstep$steps)
MDstep
```

##What is the average daily activity pattern?
calculate average steps by interval
```{r}
Avgstep <- with(mydata1,aggregate(list(steps=steps), by=list(interval=interval), mean))
head(Avgstep)
```
plot the averge steps by interval 
```{r}
g <- qplot(interval,steps,data=Avgstep,geom="line")
g+xlab("Interval")+ylab("Average Number of Steps")+ggtitle("Avergae Number of Steps by 5-minutes Interval")
```
find the interval for maxnum step
```{r}
mm <- Avgstep[which.max(Avgstep$steps),]
mm
```
##Imputing missing values
calculate total number of missing steps
```{r}
miss <- mydata[is.na(mydata$steps),]
sum(is.na(miss$steps))
```
using the average steps by interval which I calculated above to impute the missing steps
```{r}
miss$steps <- Avgstep$steps[Avgstep$interval %in% miss$interval]
```
create a new data set with imputed missing values
```{r}
mydata2 <- rbind(mydata1,miss)
```
re-calculate total number of steps by date
```{r}
TstepImput <- with(mydata2,aggregate(list(steps=steps), by=list(date=date), sum))
```
plot histogram for total steps by date again, this time use imputed data
```{r}
g <- qplot(steps, data=TstepImput, geom="histogram",bins=20)
g+xlab("Total Steps")+ylab("Frequency")+ggtitle("Total Number of Steps Taken Each Day for Imputed Steps")
```
find step mean and median
```{r}
##mean
meanImput <- mean(TstepImput$steps)
meanImput
##median
medianImput <- median(TstepImput$steps)
medianImput
```
##Are there differences in activity patterns between weekdays and weekends?
create a new variable (week) flags weekday and weekend
```{r}
mydata2$week <- ifelse((weekdays(as.Date(mydata2$date))=="Saturday" | weekdays(as.Date(mydata2$date))=="Sunday"),"Weekend","Weekday")
```
calculate mean of steps by interval and week variable
```{r}
TWstep <- with(mydata2,aggregate(list(steps=steps), by=list(interval=interval,week=week), mean))
```
plot the mean steps by interval for both weekday and weekend
```{r}
g <- qplot(interval,steps,data=TWstep,geom="line",color=week)
g+xlab("Interval")+ylab("Average Number of Steps")+ggtitle("Average Steps by Interval and by Weekday or Weekend")+facet_grid(week~.)
```
Yes, the patterns are different between weekday and weekend.